OptiFit GLNE Exploratory

### 2022-02-21 exploratory/exploratory.Rmd
updated exploratory

-   [Project Summary](#summary)
-   [Overview of data splits](#splits)
-   [Percent of reads mapped with optifit](#pct_mapped)
-   [MCC Optifit vs OptiClust](#mcc)
-   [Percent of reads mapped vs MCC](#pct_v_mcc_ref)
-   [Hyperparameter Performance](#hp_performance)
-   [Performance during training](#cv_auc)
-   [Performance on test data](#test_auc)
-   [Average ROC curves](#avg_roc)
-   [Correct Classification Frequency](#class_correct_freq)

@@

@@

### Percent Mapped VS MCC (without reference) {#pct_v_mcc_ref}
```{r pct_map_v_mcc_ref}
# 
# sub_pct <- pct_mapped %>% 
#   select(split,fraction_mapped)
# 
# mcc_pct <- merged_mcc %>% 
#   filter(algorithm == "optifit" & subset == "test",state == "fit") %>% 
#   select(split,mcc) %>% 
#   inner_join(., sub_pct ,by="split") 
# 
# mcc_pct %>% 
#   ggplot(aes(x=fraction_mapped,y=mcc)) +
#     geom_point(size=2,alpha = 0.5) +
#     theme_bw() +
#     xlab("Percent of Reads Mapped with OptiFit") +
#     ylab("OptiFit MCC (printref=f)") +
#     geom_text(aes(fraction_mapped,mcc,label=split),
#               data=subset(mcc_pct, (fraction_mapped > 0.997 | fraction_mapped < 0.98)),
#               color="black",vjust=-1,size=3) +
#     geom_text(aes(fraction_mapped,mcc,label=split),
#               data=subset(mcc_pct, (mcc > 0.775 | mcc < 0.7)),
#               color="black",vjust=-1,size=3)
```

### Percent Mapped VS MCC (WITH reference) {#pct_v_mcc_noref}
```{r pct_map_v_mcc}

@@

@@

# plot all levels together
avg_sens %>%
  ggplot(aes(x=fpr,y=mean_sensitivity,
             ymin=lower_sens,ymax=upper_sens)) +
  geom_line(aes(color=algorithm),alpha=0.7,lwd=1) +
  geom_ribbon(aes(fill=algorithm),alpha=0.3) +
  coord_equal() +
  geom_abline(intercept = 0,lty="dashed",color="grey50") +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  xlab("False Positive Rate") +
  ylab("Average True Positive Rate") +
  #facet_wrap(~level,ncol=3) +
  #scale_x_reverse(expand = c(0,0),labels = scales::percent_format(accuracy = 1)) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text = element_text(size=11),
        axis.title = element_text(size=12),
        axis.text.x = element_text(vjust = -0.5),
        axis.title.x = element_text(vjust= -0.5)) +
  guides(color = guide_legend(nrow = 1,title=""),
         fill = guide_legend(nrow = 1,title="")) +
  scale_color_manual(values=c("brown3","deepskyblue3"))

allAUCs %>% 
  group_by(algorithm) %>% 
  summarize(median_AUC = round(median(AUC),3),
            mean_AUC = round(mean(AUC),3)) %>% 
  kableExtra::kable(col.names = c("Algorithm","Median AUC","Mean AUC"))
```

### Prediction Probabilities

@@

```{r pct_correct_point}
#order samples by difference in pct_correct between algorithms
samp_order <- pct_correct %>% 
  pivot_wider(id_cols = c("Group","dx"),
              names_from = algorithm,
              values_from = pct_correct) %>% 
  mutate(difference = abs(opticlust - optifit)) %>% 
  arrange(difference) %>% 
  pull(Group)

pct_correct %>% 
  ggplot(aes(x=factor(Group,levels = samp_order),y=pct_correct,color=algorithm)) +
  geom_point(alpha=0.7) +
  facet_wrap(.~dx,nrow=2) +
  theme_bw() + 
  ylab("Percent of splits sample is correctly classified (TP or TN)") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  xlab("Sample")
```

```{r mod_pct_correct_point}
#get rid of samples that are perfectly classified with both algorithms
#and samples that have the samp pct_correct
#order samples by difference in pct_correct between algorithms
samp_order <- pct_correct %>% 
  pivot_wider(id_cols = c("Group","dx"),
              names_from = algorithm,
              values_from = pct_correct) %>% 
  filter(optifit != 1 & opticlust != 1) %>% 
  mutate(difference = abs(opticlust - optifit)) %>% 
  filter(difference > 0) %>% 
  arrange(difference) %>% 
  pull(Group)

pct_correct %>% 
  filter(Group %in% samp_order) %>% 
  ggplot(aes(x=factor(Group,levels = samp_order),y=pct_correct,color=algorithm)) +
  geom_point(alpha=0.7) +
  facet_wrap(.~dx,nrow=2) +
  theme_bw() + 
  ylab("Percent of splits sample is correctly classified (TP or TN)") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  xlab("Sample")

>>>>

-   [Project Summary](#summary)
-   [Overview of data splits](#splits)
-   [MCC Optifit vs OptiClust](#mcc)
-   [Hyperparameter Performance](#hp_performance)
-   [Performance during training](#cv_auc)
-   [Performance on test data](#test_auc)
-   [Average ROC curves](#avg_roc)
-   [Sensitivity at Specificity Values](#sens_xspec)
-   [Correct Classification Frequency](#class_correct_freq)
-   [Prediction Probabilities](#pred_prob)
-   [MCC vs Sensitivity/Specificity](#mcc_senspec)

@@

How often do samples appear in the test set together?

```{r split_together_freq}

split_count <- read_csv("../results/tables/splitTogetherFrequency.csv",col_types = cols(Group=col_character(),
                                                                                        .default = col_double()))
quant_overlap <- function(table,group1,group2){
  counts <- table %>% 
    filter(Group == c(group1,group2)) %>% 
    select(-Group) %>% 
    colSums()

  bothTest = length(which(counts == 2))
  bothTrain = length(which(counts == 0))

  tibble(group1 = group1,
         group2 = group2,
         bothTest = bothTest,
         bothTrain = bothTrain)
}

samps <- split_count %>% pull(Group)

# togetherFreq <- expand_grid(x=1:length(samps),
#             y=1:length(samps)) %>% 
#   filter(x < y) %>% 
#   mutate(group1 = samps[x],
#          group2 = samps[y]) %>% 
#   select(-x,-y) %>% 
#   group_by(group1,group2) %>% 
#   summarize(quant_overlap(split_count,group1,group2),
#             .groups="drop")
# write_csv(togetherFreq,"./data/togetherFreq.csv")
togetherFreq <- read_csv("./data/togetherFreq.csv",col_types = cols(group1=col_character(),
                                                                    group2=col_character(),
                                                                    .default=col_double()))

meanBothTest = round(mean(togetherFreq$bothTest))
meanBothTrain = round(mean(togetherFreq$bothTrain))

neverBothTest = togetherFreq %>% 
  filter(bothTest == 0) %>% 
  nrow()
neverBothTestFreq = round(neverBothTest / nrow(togetherFreq),digits=2)*100

alwaysBothTrain = togetherFreq %>% 
  filter(bothTrain == 100) %>% 
  nrow()
alwaysBothTrainFreq = round(alwaysBothTrain / nrow(togetherFreq),digits=2)*100

maxBothTest = togetherFreq %>% 
  filter(bothTest == max(bothTest))
```

For each pair of samples, I quantified how many splits both samples are in the test set together. On average, samples are both in the test set for `r meanBothTest` split. A majority of sample pairs are never in the test set together (`r neverBothTestFreq`%). At most there are `r nrow(maxBothTest)` sample pairs that are both in the test set for `r maxBothTest %>% pull(bothTest) %>% unique()` splits.

```{r together_freq_plot}
# togetherFreq %>% 
#   pivot_longer(!c(group1,group2),names_to="type",values_to = "freq") %>% 
#   ggplot(aes(x=freq,fill=type)) +
#     geom_histogram(binwidth = 1) +
#     facet_wrap(.~type,nrow=2,scales="free_y") +
#     theme_bw() +
#     scale_y_continuous(expand=c(0,0)) + 
#     scale_x_continuous(breaks = seq(0,100,10),expand=c(0,0))
# 
# meanBothTest = round(mean(togetherFreq$bothTest))
# meanBothTrain = round(mean(togetherFreq$bothTrain))

# togetherFreq %>% 
#   pivot_longer(!c(group1,group2),names_to="type",values_to = "freq") %>% 
#   ggplot(aes(x=freq,fill=type)) +
#     geom_density() +
#     facet_wrap(.~type,nrow=2,scales = "free_y") +
#     theme_bw() +
#     scale_y_continuous(expand=c(0,0)) + 
#     scale_x_continuous(breaks = seq(0,100,10),expand=c(0,0))

togetherFreq %>% 
  ggplot(aes(x=bothTest)) +
    geom_histogram(binwidth=1,fill="#F8766D",color="black",alpha=0.8) +
    theme_bw() +
    scale_y_continuous(expand=c(0,0)) +
    scale_x_continuous(expand=c(0,0)) +
    xlab("Number of splits where sample pair are both in the test set") +
    ylab("Number of Sample Pairs")

I also split the pairs by diagnosis (both cancer, both normal, or mixed) to make sure there was no difference by dx.

```{r together_freq_plot2}
sample_dx <- read_csv("../data/learning/summary/merged_predictions.csv",
                      col_types = cols(cancer=col_double(),                              
                                       normal=col_double(),
                                       .default=col_character())) %>% 
  select(Group,dx) %>% 
  distinct()

togetherFreq %>% 
  inner_join(.,sample_dx,by=c("group1" = "Group")) %>% 
  rename(group1_dx = dx) %>% 
  inner_join(.,sample_dx,by=c("group2" = "Group")) %>% 
  rename(group2_dx = dx) %>% 
  select(group1,group1_dx,group2,group2_dx,bothTest,bothTrain) %>% 
  mutate(dx=case_when(group1_dx == "normal" & group2_dx == "normal" ~ "normal",
                      group1_dx == "cancer" & group2_dx == "cancer" ~ "cancer",
                      TRUE ~ "mix")) %>% 
  ggplot(aes(x=bothTest,fill=dx)) +
    geom_histogram(binwidth=1,color="black",alpha=0.8) +
    #geom_density(color="black",alpha=0.5) +
    theme_bw() +
    scale_y_continuous(expand=c(0,0)) +
    scale_x_continuous(expand=c(0,0)) +
    xlab("Number of splits where sample pair are both in the test set") +
    ylab("Number of Sample Pairs") +
    facet_wrap(.~dx,nrow=3,scales="free_y")

# togetherFreq %>% 
#   inner_join(.,sample_dx,by=c("group1" = "Group")) %>% 
#   rename(group1_dx = dx) %>% 
#   inner_join(.,sample_dx,by=c("group2" = "Group")) %>% 
#   rename(group2_dx = dx) %>% 
#   select(group1,group1_dx,group2,group2_dx,bothTest,bothTrain) %>% 
#   mutate(dx=case_when(group1_dx == "normal" & group2_dx == "normal" ~ "normal",
#                       group1_dx == "cancer" & group2_dx == "cancer" ~ "cancer",
#                       TRUE ~ "mix")) %>% 
#   ggplot(aes(x=dx,y=bothTest,color=dx)) +
#     geom_jitter(width=0.3,height=0,alpha=0.7)
```

@@

```{r pct_map_v_mcc_ref}
### Percent Mapped VS MCC (without reference) {#pct_v_mcc_ref}
# 
# sub_pct <- pct_mapped %>% 
#   select(split,fraction_mapped)
# 
# mcc_pct <- merged_mcc %>% 
#   filter(algorithm == "optifit" & subset == "test",state == "fit") %>% 
#   select(split,mcc) %>% 
#   inner_join(., sub_pct ,by="split") 
# 
# mcc_pct %>% 
#   ggplot(aes(x=fraction_mapped,y=mcc)) +
#     geom_point(size=2,alpha = 0.5) +
#     theme_bw() +
#     xlab("Percent of Reads Mapped with OptiFit") +
#     ylab("OptiFit MCC (printref=f)") +
#     geom_text(aes(fraction_mapped,mcc,label=split),
#               data=subset(mcc_pct, (fraction_mapped > 0.997 | fraction_mapped < 0.98)),
#               color="black",vjust=-1,size=3) +
#     geom_text(aes(fraction_mapped,mcc,label=split),
#               data=subset(mcc_pct, (mcc > 0.775 | mcc < 0.7)),
#               color="black",vjust=-1,size=3)
```

```{r pct_map_v_mcc}
### Percent Mapped VS MCC (WITH reference) {#pct_v_mcc_noref}

@@

```{r allAUC}

allAUCs <- read_csv("../data/learning/summary/merged_performance.csv",
                    col_types = cols(algorithm=col_character(),
                                     method=col_character(),
                                     split=col_character(),
                                     .default=col_double())) 
meds <- allAUCs %>% 
  select(cv_metric_AUC,AUC,algorithm) %>% 
  rename(CV=cv_metric_AUC,
         Test=AUC) %>% 
  pivot_longer(!algorithm,names_to="type",values_to = "AUC") %>% 
  group_by(algorithm,type) %>% 
  summarise(median_AUC = round(median(AUC),digits=3))

allAUCs %>% 
  select(cv_metric_AUC,AUC,algorithm) %>% 
  rename(CV=cv_metric_AUC,
         Test=AUC) %>% 
  pivot_longer(!algorithm,names_to="type",values_to = "AUC") %>% 
  ggplot(aes(x=type,y=AUC,color=algorithm)) +
    geom_jitter(height=0,width=0.2,alpha = 0.6,size=2) + 
    stat_summary(fun.data=median_hilow, fun.args=0.5, geom="crossbar", width=0.5, color="black") +
    theme_bw() +
    ylab("AUC") +
    xlab("") +
    facet_grid(.~algorithm) +
    theme(legend.position = "top") +
    geom_text(data=meds,aes(x=type,y=median_AUC,label=median_AUC),
              color="black",hjust=-1.2)

@@

#calculate mean/median
mean_med <- allAUCs %>% 
  group_by(algorithm) %>% 
  summarize(median_AUC = round(median(AUC),3),
            mean_AUC = round(mean(AUC),3))

of_mean <- mean_med %>% filter(algorithm=="optifit") %>% pull(mean_AUC)
oc_mean <- mean_med %>% filter(algorithm=="opticlust") %>% pull(mean_AUC)

# plot all levels together
avg_sens %>%
  ggplot(aes(x=fpr,y=mean_sensitivity,
             ymin=lower_sens,ymax=upper_sens)) +
  geom_line(aes(color=algorithm),alpha=0.7,lwd=1) +
  geom_ribbon(aes(fill=algorithm),alpha=0.3) +
  coord_equal() +
  geom_abline(intercept = 0,lty="dashed",color="grey50") +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  xlab("False Positive Rate") +
  ylab("Average True Positive Rate") +
  #facet_wrap(~level,ncol=3) +
  #scale_x_reverse(expand = c(0,0),labels = scales::percent_format(accuracy = 1)) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text = element_text(size=11),
        axis.title = element_text(size=12),
        axis.text.x = element_text(vjust = -0.5),
        axis.title.x = element_text(vjust= -0.5)) +
  guides(color = guide_legend(nrow = 1,title=""),
         fill = guide_legend(nrow = 1,title="")) +
  scale_color_manual(values=c("brown3","deepskyblue3")) +
  geom_text(x=0.7,y=0.2,label=paste0("OptiClust Mean AUC: ",oc_mean),size=3.5) +
  geom_text(x=0.7,y=0.15,label=paste0("OptiFit Mean AUC: ",of_mean),size=3.5)

### Sensitivity at Specificity Thresholds {#sens_xspec}
```{r sens_xspec}

fun_median <- function(x){
  return(data.frame(y=round(median(x),digits=3),label=round(median(x,na.rm=T),digits=3)))
}

spec = 0.9
all_sens_spec %>% 
  filter(specificity == spec) %>% 
  ggplot(aes(x=algorithm,y=sensitivity,color=algorithm)) +
    geom_jitter(height=0,width=0.2,alpha = 0.7,size=2) +
    stat_summary(fun.data=median_hilow, fun.args=0.5, geom="crossbar", width=0.5, color="black") +
    theme_bw() +
    ylab("AUC") +
    xlab("") +
    theme(legend.position = "top") +
    ggtitle(paste0("Sensitivity at ",spec*100,"% Specificity")) +
    stat_summary(fun.data=fun_median, geom="text",show.legend = FALSE,color="black",hjust=-2.2)

spec = 0.5
all_sens_spec %>% 
  filter(specificity == spec) %>% 
  ggplot(aes(x=algorithm,y=sensitivity,color=algorithm)) +
    geom_jitter(height=0,width=0.2,alpha = 0.7,size=2) +
    stat_summary(fun.data=median_hilow, fun.args=0.5, geom="crossbar", width=0.5, color="black") +
    theme_bw() +
    ylab("AUC") +
    xlab("") +
    theme(legend.position = "top") +
    ggtitle(paste0("Sensitivity at ",spec*100,"% Specificity")) +
    stat_summary(fun.data=fun_median, geom="text",show.legend = FALSE,color="black",hjust=-2.2)

spec = 0.1
all_sens_spec %>% 
  filter(specificity == spec) %>% 
  ggplot(aes(x=algorithm,y=sensitivity,color=algorithm)) +
    geom_jitter(height=0,width=0.2,alpha = 0.7,size=2) +
    stat_summary(fun.data=median_hilow, fun.args=0.5, geom="crossbar", width=0.5, color="black") +
    theme_bw() +
    ylab("AUC") +
    xlab("") +
    theme(legend.position = "top") +
    ggtitle(paste0("Sensitivity at ",spec*100,"% Specificity")) +
    stat_summary(fun.data=fun_median, geom="text",show.legend = FALSE,color="black",hjust=-2.2)
```

```{r pct_correct_point,fig.width=11}
#order samples by difference in pct_correct between algorithms
samp_order <- pct_correct %>% 
  pivot_wider(id_cols = c("Group","dx"),
              names_from = algorithm,
              values_from = pct_correct) %>% 
  mutate(difference = abs(opticlust - optifit)) %>% 
  arrange(difference) %>% 
  pull(Group)

pct_correct %>% 
  ggplot(aes(x=factor(Group,levels = samp_order),y=pct_correct,color=algorithm)) +
  geom_point(alpha=0.7) +
  facet_wrap(.~dx,nrow=2,scales="free_x") +
  theme_bw() + 
  ylab("Percent of splits sample is correctly classified (TP or TN)") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  xlab("Sample") +
  theme(legend.position = "top")
```

```{r mod_pct_correct_point,fig.width=9}
#get rid of samples that are perfectly classified with both algorithms
#and samples that have the samp pct_correct
#order samples by difference in pct_correct between algorithms
samp_order <- pct_correct %>% 
  pivot_wider(id_cols = c("Group","dx"),
              names_from = algorithm,
              values_from = pct_correct) %>% 
  #filter(optifit != 1 & opticlust != 1) %>% 
  mutate(difference = abs(opticlust - optifit)) %>% 
  filter(difference > 0) %>% 
  arrange(difference,optifit) %>% 
  pull(Group)

pct_correct %>% 
  filter(Group %in% samp_order) %>% 
  ggplot(aes(x=factor(Group,levels = samp_order),y=pct_correct,color=algorithm)) +
  geom_point(alpha=0.7) +
  facet_wrap(.~dx,nrow=2,scales="free_x") +
  theme_bw() + 
  ylab("Percent of splits sample is correctly classified (TP or TN)") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  xlab("Sample") +
  theme(legend.position = "top")

### Prediction Probabilities {#pred_prob}
I selected the samples where the difference in the correct classification frequency between optifit and opticlust was greater than 0.5 as shown below.

```{r predprob_subset}

samp_order <- pct_correct %>% 
  pivot_wider(id_cols = c("Group","dx"),
              names_from = algorithm,
              values_from = pct_correct) %>% 
  #filter(optifit != 1 & opticlust != 1) %>% 
  mutate(difference = abs(opticlust - optifit)) %>% 
  filter(difference > 0.5) %>% 
  arrange(difference,optifit) %>% 
  pull(Group)

pct_correct %>% 
  filter(Group %in% samp_order) %>% 
  ggplot(aes(x=factor(Group,levels = samp_order),y=pct_correct,color=algorithm)) +
  geom_point(alpha=0.7) +
  facet_wrap(.~dx,nrow=2,scales="free_x") +
  theme_bw() + 
  ylab("Percent of splits sample is correctly classified (TP or TN)") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  xlab("Sample")

For each of the `r length(samp_order)` samples in this set, I plotted the probability of cancer from each split this sample was in the test set for. The dashed line is the 0.5 probability line. Using the 0.5 threshold, a point above the line means the sample would be classified as cancer in that split, below the line normal. For the cancer samples, a point above the line is a true positive while for the normal samples a point above the line is a false positive

```{r predprob, fig.height = 10, fig.width = 9}
mergedProb <- read_csv("../data/learning/summary/merged_predictions.csv",
                       col_types = cols(cancer=col_double(),
                                        normal=col_double(),
                                        .default=col_character())) %>%
  mutate(GroupDX = paste0(Group," ",dx))

ids = as.character(samp_order)

## NORMAL
mergedProb %>%
  filter(Group %in% ids) %>%
  filter(dx == "normal") %>% 
  ggplot(aes(x=algorithm,y=cancer,color=algorithm)) +
    geom_jitter(width = 0.2,height=0,alpha=0.5,size=1.8) +
    facet_wrap(.~GroupDX) +
    theme_bw() +
    geom_hline(yintercept = 0.5,lty="dashed",color="grey") +
    ylab("Probability of SRN") + xlab("") +
    scale_y_continuous(limits=c(0,1),breaks=seq(0.2,0.8,0.2)) +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          strip.background = element_rect(fill="deepskyblue3", size=1, color="black"),
          strip.text.x = element_text(face="bold", color="white"),
          legend.position = "top")
## CANCER
mergedProb %>%
  filter(Group %in% ids) %>%
  mutate(Group=factor(Group,levels=ids)) %>% 
  filter(dx == "cancer") %>% 
  ggplot(aes(x=algorithm,y=cancer,color=algorithm)) +
    geom_jitter(width = 0.2,height=0,alpha=0.5,size=1.8) +
    facet_wrap(.~GroupDX) +
    theme_bw() +
    geom_hline(yintercept = 0.5,lty="dashed",color="grey") +
    ylab("Probability of SRN") + xlab("") +
    scale_y_continuous(limits=c(0,1),breaks=seq(0.2,0.8,0.2)) +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          strip.background = element_rect(fill="brown3", size=1, color="black"),
          strip.text.x = element_text(face="bold", color="white"),
          legend.position = "top")

For comparison, here are a subset of samples where there is not a difference in the percent correctly classified between optifit and opticlust

```{r predprob_subset2}

samp_order1 <- pct_correct %>% 
  pivot_wider(id_cols = c("Group","dx"),
              names_from = algorithm,
              values_from = pct_correct) %>% 
  filter(optifit == 1 & opticlust == 1) %>% 
  mutate(difference = abs(opticlust - optifit)) %>% 
  arrange(difference,optifit) %>% 
  sample_n(10) %>% #randomly select 5
  pull(Group)

samp_order2 <- pct_correct %>% 
  pivot_wider(id_cols = c("Group","dx"),
              names_from = algorithm,
              values_from = pct_correct) %>% 
  filter(optifit < 0.9 & opticlust < 0.9) %>% 
  filter(optifit > 0.1 & opticlust > 0.1) %>% 
  mutate(difference = abs(opticlust - optifit)) %>% 
  filter(difference == 0) %>% 
  sample_n(min(10,nrow(.))) %>% #randomly select 5
  arrange(optifit) %>% 
  pull(Group)

samp_order3 <- pct_correct %>% 
  pivot_wider(id_cols = c("Group","dx"),
              names_from = algorithm,
              values_from = pct_correct) %>% 
  filter(optifit == 0 & opticlust == 0) %>% 
  mutate(difference = abs(opticlust - optifit)) %>% 
  arrange(difference,optifit) %>% 
  sample_n(10) %>% #randomly select 5
  pull(Group)

samp_order <- c(samp_order1,samp_order2,samp_order3)

pct_correct %>% 
  filter(Group %in% samp_order) %>% 
  ggplot(aes(x=factor(Group,levels = samp_order),y=pct_correct,color=algorithm)) +
  geom_point(alpha=0.7) +
  facet_wrap(.~dx,nrow=2,scales="free_x") +
  theme_bw() + 
  ylab("Percent of splits sample is correctly classified (TP or TN)") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  xlab("Sample")

```{r predprob2, fig.height = 10, fig.width = 9}
mergedProb <- read_csv("../data/learning/summary/merged_predictions.csv",
                       col_types = cols(cancer=col_double(),
                                        normal=col_double(),
                                        .default=col_character())) %>%
  mutate(GroupDX = paste0(Group," ",dx))

ids = as.character(samp_order)

## NORMAL
mergedProb %>%
  filter(Group %in% ids) %>%
  filter(dx == "normal") %>% 
  ggplot(aes(x=algorithm,y=cancer,color=algorithm)) +
    geom_jitter(width = 0.2,height=0,alpha=0.5,size=1.8) +
    facet_wrap(.~GroupDX) +
    theme_bw() +
    geom_hline(yintercept = 0.5,lty="dashed",color="grey") +
    ylab("Probability of SRN") + xlab("") +
    scale_y_continuous(limits=c(0,1),breaks=seq(0.2,0.8,0.2)) +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          strip.background = element_rect(fill="deepskyblue3", size=1, color="black"),
          strip.text.x = element_text(face="bold", color="white"),
          legend.position = "top")
## CANCER
mergedProb %>%
  filter(Group %in% ids) %>%
  mutate(Group=factor(Group,levels=ids)) %>% 
  filter(dx == "cancer") %>% 
  ggplot(aes(x=algorithm,y=cancer,color=algorithm)) +
    geom_jitter(width = 0.2,height=0,alpha=0.5,size=1.8) +
    facet_wrap(.~GroupDX) +
    theme_bw() +
    geom_hline(yintercept = 0.5,lty="dashed",color="grey") +
    ylab("Probability of SRN") + xlab("") +
    scale_y_continuous(limits=c(0,1),breaks=seq(0.2,0.8,0.2)) +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          strip.background = element_rect(fill="brown3", size=1, color="black"),
          strip.text.x = element_text(face="bold", color="white"),
          legend.position = "top")

## Does mcc correlate with sensitivity/specificity? {#mcc_senspec}

OptiFit MCC values from just the fit data compared to sensitivity/specificity (threshold = 0.5)

```{r mcc_sensspec}
library(caret)
library(assertthat)

get_sensspec_thrsh <- function(senspec,threshold){
  assert_that(is.numeric(threshold))

  senspec %>%
    mutate(prediction = case_when(cancer >= threshold ~ "cancer",
                                  TRUE ~ "normal")) %>%
    select(-cancer,-normal) %>% 
    group_by(split,algorithm) %>% 
    summarize()

  senspec %>%
    mutate(prediction = case_when(cancer >= 0.5 ~ "cancer",
                                  TRUE ~ "normal")) %>%
    mutate(dx = factor(dx,levels=c("normal","cancer")),
           prediction = factor(prediction,levels=c("normal","cancer"))) %>% 
    select(-cancer,-normal) %>% 
    mutate(category = case_when(dx == "normal" & prediction == "normal" ~ "TN",
                                dx == "cancer" & prediction == "cancer" ~ "TP",
                                dx == "cancer" & prediction == "normal" ~ "FN",
                                dx == "normal" & prediction == "cancer" ~ "FP")) %>% 
    group_by(split,algorithm) %>% 
    count(category) %>% 
    pivot_wider(id_cols=c("split","algorithm"),names_from = category,values_from = n) %>% 
    group_by(split,algorithm) %>% 
    summarise(sensitivity = TP / (TP+FN),
              specificity = TN / (TN+FP))

}

senspec_50 <- get_sensspec_thrsh(read_csv("../data/learning/summary/merged_predictions.csv"),0.5)

# senspec_50 %>% 
#   filter(algorithm == "optifit")

merged_mcc %>% 
  filter(label=="OptiFit Test" & state=="fit") %>% 
  select(-subset,-Group) %>% 
  inner_join(senspec_50 %>% filter(algorithm == "optifit"),by=c("split","algorithm")) %>% 
  ggplot(aes(x=mcc,y=sensitivity)) +
    geom_point() +
    theme_bw()

merged_mcc %>% 
  filter(label=="OptiFit Test" & state=="fit") %>% 
  select(-subset,-Group) %>% 
  inner_join(senspec_50 %>% filter(algorithm == "optifit"),by=c("split","algorithm")) %>% 
  ggplot(aes(x=mcc,y=specificity)) +
    geom_point() +
    theme_bw()

@@
